---
title: "Heart Disease Prediction"
output: html_notebook
---


```{r}
library(ggplot2)
library(PerformanceAnalytics)
library(dplyr)
library(caTools)
library(rpart)
library(rpart.plot)
library(caret)
```

# Import Data
```{r}
data <- read.csv('framingham.csv')
head(data)
summary(data)
```
## Remove Incomplete Rows
```{r}
data <- data[complete.cases(data), ]
summary(data)
```

# Exploratory Data Analysis
## Factor Distribution & Independence
```{r}

colSums(is.na(data))
for(i in 1:ncol(data)) {
  hist(data[,i], main = colnames(data)[i])
}
```
```{r}
dataNum = data[c("age", "cigsPerDay", "totChol", "sysBP", "diaBP", "BMI", "heartRate", "glucose")]
chart.Correlation(data, method="pearson", histogram=TRUE, pch=16)

```
## Removing Dependent Factors
```{r}
data <- data %>%
  select(-c("currentSmoker", "prevalentHyp", "diabetes", "diaBP"))
head(data)
```


## Outlier Cleaning
There are some numeric factors that have some outliers: cigsPerDay, totChol, sysBp,
diaBP, BMI, heartRate, glucose. We should filter out those cases.
```{r}
table(data$TenYearCHD)
data <- data %>%
  filter(cigsPerDay<=35 & 
           totChol<=420 & 
           sysBP<220 & 
           BMI<=40 & 
           heartRate<=110 & 
           glucose<=150)
table(data$TenYearCHD)
```
## Reassign variables from integers to factors
```{r}
var_factors <- c('male', 
                 'prevalentStroke',
                 'education',
                 'BPMeds',
                 'TenYearCHD')

data[var_factors] <- lapply(data[var_factors], FUN=as.factor)
colnames(data)[1] <- "gender"
levels(data$gender) <- c('female', 'male') 
levels(data$prevalentStroke) <- c('no stroke', 'stroke') 
levels(data$TenYearCHD) <- c('no CHD', 'CHD') 


```

## One Factor Analysis
```{r}
ggplot(data, aes(fill=as.factor(TenYearCHD), x=glucose)) + 
  geom_histogram(color="gray90")
```
```{r}
ggplot(data, aes(fill=TenYearCHD, x=cigsPerDay)) + 
  geom_histogram(color="gray90", bins=4) +
  facet_wrap(~gender)
```

```{r}
data %>%
  ggplot(aes(glucose, fill = TenYearCHD)) +
  geom_histogram(position = 'fill', bins = 5) + 
  facet_wrap(~ gender) 

```
```{r}
data %>%
  ggplot(aes(prevalentStroke, fill=TenYearCHD)) + 
  geom_bar(position = "fill") #position "fill" enforces proportions
```


```{r}
data %>%
  ggplot(aes(prevalentStroke, fill=TenYearCHD)) + 
  geom_bar(position = "fill") +
facet_wrap(~ gender) 
```
```{r}
data %>%
  ggplot(aes(sysBP, fill=TenYearCHD)) + 
  geom_histogram(position = "fill", bins=10) #position "fill" enforces proportions
```
```{r}
data %>%
  ggplot(aes(sysBP, fill=TenYearCHD)) + 
  geom_histogram(position = "fill", bins=5) +
  facet_wrap(~glucose>140)
```
```{r}
data %>%
  ggplot(aes(age, fill=TenYearCHD)) + 
  geom_histogram(bins=5) +
  facet_wrap(~ gender)
```
```{r}
data %>%
  ggplot(aes(totChol, fill=TenYearCHD)) + 
  geom_histogram(bins=15) +
  facet_wrap(~ gender)
```
```{r}
data %>%
  ggplot(aes(age, TenYearCHD, fill=TenYearCHD)) + 
  geom_boxplot() +
  facet_wrap(~ gender) +
  coord_flip()
```
```{r}
data %>%
  ggplot(aes(sysBP, TenYearCHD, fill=TenYearCHD)) + 
  geom_boxplot() +
  facet_wrap(~ gender) +
  coord_flip()
```

# Fitting Models
## Sampling
```{r}

set.seed(42)
split <- sample.split(Y=data$TenYearCHD, SplitRatio = 0.7) #boolean vector, not indexes
train <- data[split,]
test <- data[!split,]
```
```{r}
table(train$TenYearCHD)[1]/nrow(train)
table(test$TenYearCHD)[1]/nrow(test)
```
```{r}
set.seed(99)
split <- sample.split(Y=data$TenYearCHD, SplitRatio = 0.7) #boolean vector, not indexes
train <- data[split,]
test <- data[!split,]
train <- upSample(train, train$TenYearCHD, list=F)
train <- train[,1:ncol(train)-1]

```
```{r}
table(train$TenYearCHD)[1]/nrow(train)
table(test$TenYearCHD)[1]/nrow(test)
```

## Logistic

### One Factor Regression
```{r}
modelAge = glm(TenYearCHD~age,data=train,family='binomial')
summary(modelAge)
```
```{r}
modelBP = glm(TenYearCHD~sysBP,data=train,family='binomial')
summary(modelBP)
```
### All Factor Regression
```{r}
modelLogistic = glm(TenYearCHD~.,data=train,family='binomial', control = list(maxit = 50))
summary(modelLogistic)
cutoff <- 0.53
```
```{r}
predLogisticTrain <- predict(modelLogistic, type='response', train)
predLogisticTrainBinary <- ifelse(predLogisticTrain > cutoff, 1, 0)
confusion_matrix <- table(what_happened = train$TenYearCHD, model_predicts = predLogisticTrainBinary)
# confusion_matrix
accuracy <- sum(confusion_matrix[1,1],confusion_matrix[2,2])/nrow(train)
precision <- confusion_matrix[2,2]/sum(confusion_matrix[2,2],confusion_matrix[1,2])
recall <- confusion_matrix[2,2]/sum(confusion_matrix[2,2], confusion_matrix[2,1])
F1 = 2*((recall * precision) / (recall + precision))

print("Accuracy"); print(sprintf("%.3f", accuracy));
print("Precision"); print(sprintf("%.3f", precision))
print("Recall"); print(sprintf("%.3f", recall))
print("F1"); print(sprintf("%.3f", F1))
```
```{r}
predLogisticTest <- predict(modelLogistic, type='response', test)
predLogisticTestBinary <- ifelse(predLogisticTest > cutoff, 1, 0)
confusion_matrix <- table(what_happened = test$TenYearCHD, model_predicts = predLogisticTestBinary)
#confusion_matrix
accuracy <- sum(confusion_matrix[1,1],confusion_matrix[2,2])/nrow(test)
precision <- confusion_matrix[2,2]/sum(confusion_matrix[2,2],confusion_matrix[1,2])
recall <- confusion_matrix[2,2]/sum(confusion_matrix[2,2], confusion_matrix[2,1])
F1 = 2*((recall * precision) / (recall + precision))

print("Accuracy"); print(sprintf("%.3f", accuracy)); 
print("Precision"); print(sprintf("%.3f", precision))
print("Recall"); print(sprintf("%.3f", recall))
print("F1"); print(sprintf("%.3f", F1))
```


## Trees

### Decision Tree
```{r}
set.seed(22)
tree <- rpart(TenYearCHD~sysBP+gender+age+education+cigsPerDay+BPMeds+prevalentStroke+totChol+BMI+heartRate+glucose, data=train, method='class', cp=0.015, minsplit=20)

rpart.plot(tree, varlen = 10 , tweak = 1.2, extra='auto', box.palette = "auto", type=2)
```

